---
title: "Visualizations_Drafts"
output: html_document
---

# NBA Data Visualizations
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(ggrepel)
library(forcats)
library(RColorBrewer)
library(scales)
library(ggridges)
library(patchwork)
```



# Read the data
```{r}
nba_data <- read.csv("nba_data_processed.csv")
```

# Custom color palette for your visualizations
```{r}
nba_palette <- c("#003399", "#E31837", "#007A33", "#552583", "#CE1141", 
                "#98002E", "#00538C", "#0E2240", "#C8102E", "#1D1160",
                "#6F2DA8", "#00A4BD", "#FF7F00", "#B10DC9", "#FF851B", 
                "#39CCCC", "#01FF70", "#85144b", "#F012BE", "#3D9970")
```



# 1. Points vs Minutes Played with Position Highlighted
```{r}
ggplot(nba_data, aes(x = MP, y = PTS, color = Pos)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = nba_palette) +
  labs(title = "Relationship Between Minutes Played and Points Scored",
       subtitle = "By Position",
       x = "Minutes Played",
       y = "Total Points",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")
```


# 2. Field Goal Percentage Distribution by Position
```{r}
ggplot(nba_data, aes(x = Pos, y = "FG%", fill = Pos)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, alpha = 0.5) +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Field Goal Percentage Distribution by Position",
       x = "Position",
       y = "Field Goal Percentage",
       fill = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")
```


# 3. Top 20 Scorers - Horizontal Bar Chart
```{r}
nba_data$score_rank <- rank(-nba_data$PTS)
nba_top_scorers <- nba_data[nba_data$score_rank <= 20, ]
nba_top_scorers <- nba_top_scorers[order(-nba_top_scorers$PTS), ]

ggplot(nba_top_scorers, aes(x = reorder(Player, PTS), y = PTS, fill = Tm)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Top 20 NBA Scorers",
       x = "Player",
       y = "Total Points",
       fill = "Team") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")
```


# 4. Age Distribution Density Plot with Average Markers
```{r}
ggplot(nba_data, aes(x = Age)) +
  geom_density(fill = "#1D428A", alpha = 0.7) +
  geom_vline(aes(xintercept = mean(Age)), color = "#FFC72C", linetype = "dashed", size = 1) +
  annotate("text", x = mean(nba_data$Age) + 3, y = 0.1, 
           label = paste("Mean Age:", round(mean(nba_data$Age), 1)), color = "#FFC72C") +
  labs(title = "Age Distribution of NBA Players",
       subtitle = "With Mean Age Marker",
       x = "Age",
       y = "Density") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```


# 5. Scatter Plot of Assists vs Turnovers with Size as Minutes Played
```{r}
ggplot(nba_data, aes(x = AST, y = TOV, size = MP, color = Pos)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10)) +
  scale_color_manual(values = nba_palette) +
  labs(title = "Relationship Between Assists and Turnovers",
       subtitle = "Size Represents Minutes Played",
       x = "Assists",
       y = "Turnovers",
       size = "Minutes Played",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```


# 6. Free Throw Rate (FTA/FGA) by Position
```{r}
nba_data$FT_Rate <- nba_data$FTA / (nba_data$FGA + 0.001)

ft_rate_95pct <- quantile(nba_data$FT_Rate, 0.95, na.rm = TRUE)
plot_data <- nba_data[nba_data$FT_Rate <= ft_rate_95pct, ]

ggplot(plot_data, aes(x = Pos, y = FT_Rate, fill = Pos)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, alpha = 0.5) +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Free Throw Rate by Position",
       subtitle = "Free Throw Attempts per Field Goal Attempt",
       x = "Position",
       y = "Free Throw Rate (FTA/FGA)",
       fill = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")
```


# 7. Radar Chart for Top 5 Players (using Polar Coordinates)
```{r}
top_indices <- order(nba_data$PTS, decreasing = TRUE)[1:5]

top_5_players <- nba_data[top_indices, ]

player_names <- top_5_players$Player
stats_cols <- c("PTS", "AST", "TRB", "STL", "BLK")
top_5_stats <- top_5_players[, stats_cols]

normalized_stats <- matrix(0, nrow = 5, ncol = 5)
colnames(normalized_stats) <- stats_cols
rownames(normalized_stats) <- player_names

for (i in 1:5) {
  stat_col <- stats_cols[i]
  min_val <- min(top_5_stats[, stat_col])
  max_val <- max(top_5_stats[, stat_col])
  normalized_stats[, i] <- (top_5_stats[, stat_col] - min_val) / (max_val - min_val)
}

radar_data <- data.frame(
  Player = rep(player_names, each = 5),
  Stat = rep(stats_cols, times = 5),
  Value = as.vector(t(normalized_stats))
)

ggplot(radar_data, aes(x = Stat, y = Value, group = Player, color = Player)) +
  geom_polygon(fill = NA, size = 1) +
  geom_point(size = 3) +
  coord_polar() +
  scale_color_manual(values = nba_palette[1:5]) +
  labs(title = "Performance Comparison of Top 5 Scorers",
       x = "",
       y = "Normalized Value",
       color = "Player") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_line(color = "gray80", size = 0.2))
```


# 8. Heatmap of Correlation Between Key Stats
```{r}
corr_vars <- c("PTS", "AST", "TRB", "STL", "BLK", "TOV", "PF", "MP") 

corr_data <- nba_data[, corr_vars]
corr_matrix <- cor(corr_data, use = "pairwise.complete.obs")
corr_df <- as.data.frame(as.table(corr_matrix))
names(corr_df) <- c("Var1", "Var2", "Correlation")

ggplot(corr_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "#CE1141", mid = "white", high = "#007A33", midpoint = 0) +
  geom_text(aes(label = round(Correlation, 2)), size = 2.5) +
  labs(title = "Correlation Heatmap of Key NBA Statistics",
       x = "",
       y = "",
       fill = "Correlation") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```



# 9. Age Distribution by Position using Ridge Plot
```{r}
ggplot(nba_data, aes(x = Age, y = Pos, fill = Pos)) +
  geom_density_ridges(alpha = 0.7) +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Age Distribution by Position",
       x = "Age",
       y = "Position",
       fill = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")
```


# 10. Bubble Chart for Points, Rebounds, and Assists by Position
```{r}
ggplot(nba_data, aes(x = TRB, y = AST, size = PTS, color = Pos)) +
  geom_point(alpha = 0.6) +
  scale_size_continuous(range = c(1, 15)) +
  scale_color_manual(values = nba_palette) +
  labs(title = "Points, Rebounds, and Assists by Position",
       x = "Total Rebounds",
       y = "Assists",
       size = "Points",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

# 11. Stacked Bar Chart of Offensive vs Defensive Rebounds by Position

```{r}

positions <- unique(nba_data$Pos)
orb_by_pos <- numeric(length(positions))
drb_by_pos <- numeric(length(positions))

for (i in 1:length(positions)) {
  pos <- positions[i]
  pos_data <- nba_data[nba_data$Pos == pos, ]
  orb_by_pos[i] <- sum(pos_data$ORB, na.rm = TRUE)
  drb_by_pos[i] <- sum(pos_data$DRB, na.rm = TRUE)
}


rebound_summary <- data.frame(
  Position = rep(positions, 2),
  Rebound_Type = c(rep("ORB", length(positions)), rep("DRB", length(positions))),
  Count = c(orb_by_pos, drb_by_pos)
)

rebound_summary$Position <- factor(rebound_summary$Position, levels = positions)
rebound_summary$Rebound_Type <- factor(rebound_summary$Rebound_Type, levels = c("ORB", "DRB"))


ggplot(rebound_summary, aes(x = Position, y = Count, fill = Rebound_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("ORB" = "#CE1141", "DRB" = "#007A33"),
                   labels = c("ORB" = "Offensive Rebounds", "DRB" = "Defensive Rebounds")) +
  labs(title = "Offensive vs Defensive Rebounds by Position",
       x = "Position",
       y = "Total Rebounds",
       fill = "Rebound Type") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")
```


# 12. Hexbin Plot for Field Goal Attempts vs Points
```{r}
ggplot(nba_data, aes(x = FGA, y = PTS)) +
  geom_hex(bins = 30) +
  scale_fill_viridis_c() +
  labs(title = "Relationship Between Field Goal Attempts and Points",
       subtitle = "Using Hexagonal Binning",
       x = "Field Goal Attempts",
       y = "Total Points",
       fill = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```


# 13. Scatter Plot with Efficiency Metrics
```{r}
ggplot(nba_data, aes(x = "FG%", y = PTS, color = Pos)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = nba_palette) +
  labs(title = "Field Goal Percentage vs Points Scored",
       x = "Field Goal Percentage",
       y = "Total Points",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

# 14. Minutes Played vs Age with Games Played as Size
```{r}
ggplot(nba_data, aes(x = Age, y = MP, size = G, color = Pos)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = nba_palette) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Minutes Played vs Age",
       subtitle = "Size Represents Games Played",
       x = "Age",
       y = "Minutes Played",
       size = "Games Played",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

# 15. Faceted Histogram of Points by Position
```{r}
library(ggridges)
ggplot(nba_data, aes(x = PTS, y = Pos, fill = Pos)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5, rel_min_height = 0.01) +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Distribution of Points Scored by Position",
       subtitle = "Using ridgeline plots for direct comparison",
       x = "Total Points",
       y = "") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none",
        panel.grid.major.y = element_blank())
```

# 16. Defensive and Offensive Metrics
```{r}
ggplot(nba_data, aes(x = STL, y = BLK, color = Pos, size = AST)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = nba_palette) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Defensive Metrics: Steals vs Blocks",
       subtitle = "Size Represents Assists (Offensive Contribution)",
       x = "Steals",
       y = "Blocks",
       color = "Position",
       size = "Assists") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

# 17. Bubble Plot: 3-Point Attempts, Field Goal Percentage, and Points
```{r}
ggplot(nba_data, aes(x = '3PA', y = 'FG%', size = PTS, color = Pos)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = nba_palette) +
  scale_size_continuous(range = c(1, 15)) +
  labs(title = "3-Point Attempts vs Field Goal Percentage",
       subtitle = "Size Represents Total Points",
       x = "3-Point Attempts",
       y = "Field Goal Percentage",
       color = "Position",
       size = "Points") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

# 18. Free Throw Attempts vs Field Goal Attempts by Position
```{r}
ggplot(nba_data, aes(x = FGA, y = FTA, color = Pos)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_manual(values = nba_palette) +
  labs(title = "Relationship Between Free Throw Attempts and Field Goal Attempts",
       subtitle = "By Position",
       x = "Field Goal Attempts",
       y = "Free Throw Attempts",
       color = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")
```



# 19. Minutes Played per Game by Position
```{r}
nba_data$MPG <- nba_data$MP / nba_data$G

ggplot(nba_data, aes(x = Pos, y = MPG, fill = Pos)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, alpha = 0.5) +
  scale_fill_manual(values = nba_palette) +
  labs(title = "Minutes Played per Game by Position",
       x = "Position",
       y = "Minutes per Game",
       fill = "Position") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")
```


# 20. Parallel Coordinate Plot for Top 10 Players

```{r}
top_indices <- order(nba_data$PTS, decreasing = TRUE)[1:10]
top_10_players <- nba_data[top_indices, ]

plot_data <- top_10_players[, c("Player", "PTS", "AST", "TRB", "STL", "BLK")]

library(GGally)

player_colors <- setNames(nba_palette[1:10], plot_data$Player)

ggparcoord(plot_data, 
           columns = 2:6,  
           groupColumn = "Player",
           scale = "uniminmax", 
           alphaLines = 0.8,
           showPoints = TRUE) +
  scale_color_manual(values = player_colors) +
  labs(title = "Parallel Coordinate Plot of Top 10 Players",
       x = "Statistic",
       y = "Scaled Value") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")
```